```{r, echo=FALSE,results='hide',message=FALSE,comment=""}
library(ggplot2)
```
#Reproducible Research: Peer Assessment 1  

##Loading and preprocessing the data


```{r, results="hide"}
unzip(zipfile="repdata-data-activity.zip")
data <- read.csv("activity.csv")
```

##What is mean total number of steps taken per day?  
Make a histogram of the total number of steps taken each day
```{r, results='hide',comment="",message=FALSE}
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
```

Mean total number of steps taken per day:  
```{r, results="hide"}
mean(total.steps, na.rm=TRUE)
```

Median total number of steps taken per day:
```{r, results="hide"}
median(total.steps, na.rm=TRUE)
```


##What is the average daily activity pattern?
```{r graph, results="hide",message=FALSE}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps taken")
```  
  
On average across all the days in the dataset, the 5-minute interval contains the maximum number of steps?  
```{r, results="hide"}
averages[which.max(averages$steps),]
```

##Imputing missing values

Total number of missing values:  
```{r, results="hide"}
missing <- is.na(data$steps)
table(missing)
```

Replace each missing value with the mean value of its 5-minute interval:  
```{r, results="hide"}
fill.value <- function(steps, interval) {
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else
    filled <- (averages[averages$interval==interval, "steps"])
  return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```

Histogram, Mean and Median ot the total number of steps each day with the new data set:  
```{r, results="hide"}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day with no missing data")
mean(total.steps)
median(total.steps)
```

Mean and median values were shorter before imputing missing data because missing values NA were considered as "zero" and after miputing those values are replaced by mean values.  

##Are there differences in activity patterns between weekdays and weekends?

Create a new factor varible in the dataset with two levels - "weekday" and "weekend".  
```{r, results="hide"}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)
```

Make a panel plot containing plots of average number of steps taken on weekdays and weekends.
```{r, results="hide"}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
xlab("5-minute interval") + ylab("Number of steps")
```
